\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-12-12 adaptive-estimation.tex
Finishing part (ii) of SGD theorem convergence

The starting point for our first estimator is to note that the median of a
distribution minimizes $\E[|X - \theta|]$ over $\theta \in \R$, and
moreover, we have the familiar result (cf.~\cite{VanDerVaart98}, Ch.~21)
that given a sample $X_1^n \simiid P$, if $\theta = \mbox{med}(P)$ and
$P$ has density $f$ near $\theta$, then
\begin{equation*}
  \sqrt{n}(\mbox{med}(X_1^n) - \theta)
  \cd \normal\left(0, \frac{1}{4 f(\theta)^2}\right),
\end{equation*}
which is precisely the variance lower bound in
Theorem~\ref{thm:adpative_lower_bound}.  Thus, it it is natural consider a
stochastic gradient procedure for minimizing $\E[|X - \theta|]$. To that end,
let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive
sequence of stepsizes,
and define the sequence
\begin{equation}
  \label{eq:sgd_alg}
  \theta_n = \theta_{n-1} + \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation*}
  B_n = \sgn (X_n - \theta_{n-1}).
\end{equation*}
We make one of two assumptions on the stepsizes $\gamma_n$, which
are relatively standard: we always have $\gamma_n$ non-increasing, and
\begin{subequations}
  For some $0 < \lambda \le 1$,
  \begin{align}
    \label{eqn:lazy-gamma}
    \frac{\gamma_n - \gamma_{n+1}}{\gamma_n^2}
    \to 0, & ~~~
    \sum_n \frac{\gamma_n^\frac{1 + \lambda}{2}}{\sqrt{n}} < \infty
    ~~ \mbox{or} \\
    \gamma_n = o(n^{-2/3}),
    & ~~~
    \sum_n \gamma_n = \infty.
    \label{eqn:stringent-gamma}
  \end{align}
\end{subequations}

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(\theta)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    Assume in addition that $f(x)$ is continuously differentiable with a
    finite Fisher information
    \begin{equation*}
      \sigma_f^2 \triangleq \ex{\frac{f'(X)}{f(X)} } 
    \end{equation*}
    for the location in the family $\{P_\theta\}$.
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    Then for any bounded, symmetric, and quasi-convex function $L$, 
    \begin{align} 
      & \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right], \label{eq:attaining_LAM}
    \end{align}
    where $Z \sim \Ncal(0,1)$. 
    %
    %
  \item Assume the stepsizes $\gamma_n$ satisfy both
    conditions~\eqref{eqn:lazy-gamma} and~\eqref{eqn:stringent-gamma}.
    Then
    \begin{equation*}
      \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{4 n f(\theta)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}

>>>>

The starting point for our first estimator is to note that the median of a
distribution minimizes $\E[|X - \theta|]$ over $\theta \in \R$, and
moreover, we have the familiar result (cf.~\cite{VanDerVaart98}, Ch.~21)
that given a sample $X_1^n \simiid P$, if $\theta = \mbox{med}(P)$ and
$P$ has continuous density $f(\cdot - \theta)$ near $\theta$, then
\begin{equation*}
  \sqrt{n}(\mbox{med}(X_1^n) - \theta)
  \cd \normal\left(0, \frac{1}{4 f(0)^2}\right),
\end{equation*}
which is precisely the variance lower bound in
Theorem~\ref{thm:adpative_lower_bound}.  Thus, it it is natural consider a
stochastic gradient procedure for minimizing $\E[|X - \theta|]$. To that end,
let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive
sequence of stepsizes,
and define the sequence
\begin{equation}
  \label{eq:sgd_alg}
  \theta_n = \theta_{n-1} + \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation*}
  B_n = \sgn (X_n - \theta_{n-1}).
\end{equation*}
We make one of two assumptions on the stepsizes $\gamma_n$, which
are relatively standard: we always have $\gamma_n$ non-increasing, and
\begin{subequations}
  For some $0 < \lambda \le 1$,
  \begin{align}
    \label{eqn:lazy-gamma}
    \frac{\gamma_n - \gamma_{n+1}}{\gamma_n^2}
    \to 0, & ~~~
    \sum_n \frac{\gamma_n^\frac{1 + \lambda}{2}}{\sqrt{n}} < \infty
    ~~ \mbox{or} \\
    \gamma_n = o(n^{-2/3}),
    & ~~~
    \sum_n \gamma_n = \infty.
    \label{eqn:stringent-gamma}
  \end{align}
\end{subequations}

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(\theta)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    For any bounded, symmetric, and quasi-convex function $L$, 
    \begin{align} 
      & \sup_{c < \infty} \limsup_{n \to \infty}
      \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}
      \E_\tau \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[L(W/ 2 f(0)) \right],
        \label{eq:attaining_LAM}
    \end{align}
    where $W \sim \normal(0,1)$. 
    %
    %
  \item \label{item:sgd-ms-convergence}
    Assume the stepsizes $\gamma_n$ satisfy both
    conditions~\eqref{eqn:lazy-gamma} and~\eqref{eqn:stringent-gamma}.
    Then
    \begin{equation*}
      \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{4 n f(\theta)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}
